### 2010-04-03 papers/icfp-2010-desc/paper_type_theory.tex
Desc@ICFP: definitely fix spacing of operators.

We first define the `small product' $\spi{\!}{\!}$ operator:
%
\[\stk{
%% spi
\spi{}{}: \PITEL{\V{E}}{\EnumU}\PITEL{\V{P}}{\EnumT{\V{E}} \To \Set} \To \Set \\
\begin{array}{@{}l@{\:}l@{\:\mapsto\:\:}l}
\spi{\NilE}{& \V{P}}        & \Unit \\
\spi{(\ConsE{\V{t}}{\V{E}})}{& \V{P}} & \TIMES{\V{P}\: \Ze}{\spi{\V{E}}{(\LAM{\V{x}} \V{P}\: (\Su{\V{x}}))}}
\end{array}
}\]
%
This builds a right-nested
tuple type, demanding an object of $P\:i$ for each $i$ in the given finite
domain. We can see these tuples as `jump tables' tabulating dependently
typed functions from the domain. We give this functional
interpretation---the eliminator we need---by the
$\F{switch}$ operator, which, unsurprisingly, iterates projection:
%
\[\stk{
%% switch
\begin{array}{@{}ll}
\F{switch} : \PITEL{\V{E}}{\EnumU}
               \PITEL{\V{P}}{\EnumT{\V{E}} \To \Set} \To
              \spi{\V{E}}{\V{P}} \To
               \PITEL{\V{x}}{\EnumT{\V{E}}} \To \V{P}\: \x
\end{array} \\
\begin{array}{@{}l@{\:\mapsto\:\:}l}
\switch{(\ConsE{\V{t}}{\V{E}})}{\V{P}}{\V{b}}{\Ze}      & \fst{\V{b}} \\
\switch{(\ConsE{\V{t}}{\V{E}})}{\V{P}}{\V{b}}{(\Su{\V{x}})} & \switch{\V{E}}{(\LAM{\V{x}} \V{P}
  (\Su{\V{x}}))}{(\snd{\V{b}})}{\V{x}}
\end{array}
}\]

@@

The $\spi{\!}{\!}$ and $\F{switch}$ operators deliver dependent
elimination for finite enumerations, but are rather awkward to
use directly. We do not write the range for a \(\LAMBINDER\)-abstraction, so
it is galling to supply \(\V{P}\) for functions defined by $\F{switch}$.
Let us therefore find a way to recover the tedious details of the
encoding from types.

@@

Firstly, we present the codes for enumerations as right-nested tuples
which, by our LISP convention, we write as unpunctuated lists of tags
\(\sqr{\etag{t_0}\ldots\etag{t_n}}\).
Secondly, we can denote an element \emph{by its
name}: the type pushed in allows us to recover the numerical
index. We retain the numerical forms to facilitate
\emph{generic} operations and ensure that shadowing is punished
fittingly, not fatally.
Finally, we express functions from enumerations as tuples.
Any tuple-form, \(\void\) or \(\pair{\_}{\_}{}\), is accepted by the
function space---the generalised product---if it is accepted by the
small product. Propagation fills in the appeal to $\F{switch}$,
copying the range information.

>>>>

We first define the `small product' $\SYMBspi$ operator:
%
\[\stk{
%% spi
\spi{}{}: \PITEL{\V{E}}{\EnumU}\PITEL{\V{P}}{\EnumT{\V{E}} \To \Set} \To \Set \\
\begin{array}{@{}l@{\:}l@{\:\mapsto\:\:}l}
\spi{\NilE}{& \V{P}}        & \Unit \\
\spi{(\ConsE{\V{t}}{\V{E}})}{& \V{P}} & \TIMES{\V{P}\: \Ze}{\spi{\V{E}}{(\LAM{\V{x}} \V{P}\: (\Su{\V{x}}))}}
\end{array}
}\]
%
This builds a right-nested
tuple type, demanding an object of $\V{P}\:\V{i}$ for each $\V{i}$ in the given finite
domain. We can see these tuples as `jump tables' tabulating dependently
typed functions from the domain. We give this functional
interpretation---the eliminator we need---by the
$\SYMBswitch$ operator, which, unsurprisingly, iterates projection:
%
\[\stk{
%% switch
\begin{array}{@{}ll}
\SYMBswitch : \PITEL{\V{E}}{\EnumU}
               \PITEL{\V{P}}{\EnumT{\V{E}} \To \Set} \To
              \spi{\V{E}}{\V{P}} \To
               \PITEL{\V{x}}{\EnumT{\V{E}}} \To \V{P}\: \x
\end{array} \\
\begin{array}{@{}l@{\:\mapsto\:\:}l}
\switch{(\ConsE{\V{t}}{\V{E}})}{\V{P}}{\V{b}}{\Ze}      & \fst{\V{b}} \\
\switch{(\ConsE{\V{t}}{\V{E}})}{\V{P}}{\V{b}}{(\Su{\V{x}})} & \switch{\V{E}}{(\LAM{\V{x}} \V{P}
  (\Su{\V{x}}))}{(\snd{\V{b}})}{\V{x}}
\end{array}
}\]

@@

The $\SYMBspi$ and $\SYMBswitch$ operators deliver dependent
elimination for finite enumerations, but are rather awkward to
use directly. We do not write the range for a \(\LAMBINDER\)-abstraction, so
it is galling to supply \(\V{P}\) for functions defined by $\SYMBswitch$.
Let us therefore find a way to recover the tedious details of the
encoding from types.

@@

Firstly, we present the codes for enumerations as right-nested tuples
which, by our LISP convention, we write as unpunctuated lists of tags
\(\sqr{\etag{t_0}\ldots\etag{t_n}}\).
Secondly, we can denote an element \emph{by its
name}: the type pushed in allows us to recover the numerical
index. We retain the numerical forms to facilitate
\emph{generic} operations and ensure that shadowing is punished
fittingly, not fatally.
Finally, we express functions from enumerations as tuples.
Any tuple-form, \(\void\) or \(\pair{\_}{\_}{}\), is accepted by the
function space---the generalised product---if it is accepted by the
small product. Propagation fills in the appeal to $\SYMBswitch$,
copying the range information.
